---
title: "CN Models from NEON data"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How are CN ratios related in the NEON data?

Used data set up from Adrianne and Andrew. Added litter CN from litter_chem.csv.

```{r include = FALSE}

require(tidyr)
require(dplyr)
require(ggplot2)
require(plyr) #for count function
require(plotly)
require(viridis)
require(lme4)
library(nlme)
library(car)
library(lattice)
library(grid)
library(sjPlot)

setwd('C:/Users/Anika/Documents/GradSchool/NEON_2019/')

# Load and clean data ----------------------------------------------------------------------
#load nitrogen pool data
plot.df <- read.csv(file = "CN_plotID.csv", header = T, stringsAsFactors = F)
summary(plot.df)
#load climate data
climate.df <- read.csv(file = "MAT_MAP_Allsites.csv", stringsAsFactors = F)
climate.df$MAP <- climate.df$MAP*10 # from cm to mm
#climate.df$siteID<-climate.df$SiteID
head(climate.df)

#merge the plot and climate data by site name
plot.df <- left_join(plot.df, climate.df, by = c("siteID" = "SiteID"))

#select key columns
myvars <- c('domainID', "siteID", "plotID","rootNPercent",'rootCPercent','rootCNratio',
            'rootSample_n','plotType','litterNPercent','foliarNPercent','foliarCPercent',
            'foliarCNRatio','soilNPercent_MHoriz','soilCNRatio_MHoriz','nlcdClass',
            'x','y', 'MAT', 'MAP') #need litter CN

plot.df <- plot.df[myvars]

#get litter CN ratios
litterchem <- read.csv(file = 'litter_chem.csv', header = T, stringsAsFactors = F)
litterchem <- litterchem %>% dplyr::select(plotID, siteID, CNratio)
litterchem <- litterchem %>% 
  group_by(plotID) %>% 
  dplyr::summarise(litterCNratio = mean(CNratio, na.rm=T))
litterchem <- data.frame(litterchem)
plot.df <- left_join(plot.df, litterchem, by = c("plotID" = "plotID"))

#remove duplicates resulting from the merge (linked to slight differences in coordinates?)
nrow(plot.df)
plot.df <- plot.df[!duplicated(plot.df), ]
nrow(plot.df) #check that this worked

#look at the data
unique(plot.df$nlcdClass)
unique(plot.df$siteID)

plot(plot.df$litterNPercent ~ plot.df$foliarNPercent)
### univariate distributions
hist(plot.df$foliarNPercent)
hist(plot.df$litterNPercent)
hist(plot.df$rootNPercent)
hist(log(plot.df$soilNPercent_MHoriz))
hist(plot.df$litterCNratio)
hist(plot.df$foliarCNRatio)
hist(plot.df$soilCNRatio_MHoriz)

#filter out sight (GUAN) with anamolously high soil nitrogen...
plot.df <- plot.df %>%
  dplyr::group_by(siteID) %>%
  dplyr::filter(!siteID=='GUAN')

unique(plot.df$siteID)


head(plot.df)
glimpse(plot.df)

plot.df$siteID <- as.factor(plot.df$siteID)

```

Look at CN ratios:

```{r histograms}
hist(plot.df$litterCNratio)
hist(plot.df$foliarCNRatio)
hist(plot.df$soilCNRatio_MHoriz)
```

## CN Models directly

Using foliar CN, litter CN, and soil CN as response variables:

```{r models1}
#### CN ratios directly
#litter
mod.litter1 <- lmer(litterCNratio ~ soilCNRatio_MHoriz + foliarCNRatio + (1|siteID), data=plot.df)
mod.litter2 <- lmer(litterCNratio ~ soilCNRatio_MHoriz + foliarCNRatio + soilNPercent_MHoriz + (1|siteID), data=plot.df)
mod.litter3 <- lmer(litterCNratio ~ soilCNRatio_MHoriz*MAP + foliarCNRatio + soilNPercent_MHoriz + (1|siteID), data=plot.df)

#foliar
mod.foliar1 <- lmer(foliarCNRatio ~ litterCNratio + soilCNRatio_MHoriz + (1|siteID), data=plot.df)
mod.foliar2 <- lmer(foliarCNRatio ~ litterCNratio + soilCNRatio_MHoriz + soilNPercent_MHoriz + (1|siteID), data=plot.df)
mod.foliar3 <- lmer(foliarCNRatio ~ litterCNratio + soilCNRatio_MHoriz*MAP + soilNPercent_MHoriz + (1|siteID), data=plot.df)

#soil
mod.soil1 <- lmer(soilCNRatio_MHoriz ~ litterCNratio + foliarCNRatio + (1|siteID), data=plot.df)
mod.soil2 <- lmer(soilCNRatio_MHoriz ~ litterCNratio + foliarCNRatio + soilNPercent_MHoriz + (1|siteID), data=plot.df)
mod.soil3 <- lmer(soilCNRatio_MHoriz ~ litterCNratio + foliarCNRatio*MAP + soilNPercent_MHoriz + (1|siteID), data=plot.df)

#summary comparison of models
tab_model(mod.litter1, mod.litter2, mod.litter3)
tab_model(mod.foliar1, mod.foliar2, mod.foliar3)
tab_model(mod.soil1, mod.soil2, mod.soil3)

#visualize best model from each (based on marginal and conditional R2)
sjPlot::plot_model(mod.litter3, 
                   #axis.labels=c("Urchin", "Depth", "Fish"),
                   show.values=TRUE, show.p=TRUE,
                   title="Effect of Parameters on litter CN Ratio")
sjPlot::plot_model(mod.foliar3, 
                   #axis.labels=c("Urchin", "Depth", "Fish"),
                   show.values=TRUE, show.p=TRUE,
                   title="Effect of Parameters on litter CN Ratio")
sjPlot::plot_model(mod.soil2, 
                   #axis.labels=c("Urchin", "Depth", "Fish"),
                   show.values=TRUE, show.p=TRUE,
                   title="Effect of Parameters on litter CN Ratio")
```

MAP still important here.

Delta CN between pools might be interesting to explore.

```{r delta cn}
plot.df$CNfoliar_CNlitter <- plot.df$foliarCNRatio - plot.df$litterCNratio
plot.df$CNlitter_CNsoil <- plot.df$litterCNratio - plot.df$soilCNRatio_MHoriz
plot.df$CNfoliar_CNsoil <- plot.df$foliarCNRatio - plot.df$soilCNRatio_MHoriz

hist(plot.df$CNfoliar_CNlitter)
hist(plot.df$CNlitter_CNsoil)
hist(plot.df$CNfoliar_CNsoil)

mod.foliarlitter1 <- lmer(CNfoliar_CNlitter ~ litterCNratio + soilCNRatio_MHoriz + (1|siteID), data=plot.df)
mod.foliarlitter2 <- lmer(CNfoliar_CNlitter ~ litterCNratio + soilCNRatio_MHoriz + soilNPercent_MHoriz + (1|siteID), data=plot.df)
mod.foliarlitter3 <- lmer(CNfoliar_CNlitter ~ litterCNratio + soilCNRatio_MHoriz*MAP + soilNPercent_MHoriz + (1|siteID), data=plot.df)


mod.littersoil1 <- lmer(CNlitter_CNsoil ~ soilCNRatio_MHoriz + foliarCNRatio + (1|siteID), data=plot.df)
mod.littersoil2 <- lmer(CNlitter_CNsoil ~ soilCNRatio_MHoriz + foliarCNRatio + soilNPercent_MHoriz + (1|siteID), data=plot.df)
mod.littersoil3 <- lmer(CNlitter_CNsoil ~ soilCNRatio_MHoriz*MAP + foliarCNRatio + soilNPercent_MHoriz + (1|siteID), data=plot.df)


mod.foliarsoil1 <- lmer(CNfoliar_CNsoil ~ litterCNratio + soilCNRatio_MHoriz + (1|siteID), data=plot.df)
mod.foliarsoil2 <- lmer(CNfoliar_CNsoil ~ litterCNratio + soilCNRatio_MHoriz + soilNPercent_MHoriz + (1|siteID), data=plot.df)
mod.foliarsoil3 <- lmer(CNfoliar_CNsoil ~ litterCNratio + soilCNRatio_MHoriz*MAP + soilNPercent_MHoriz + (1|siteID), data=plot.df)

#summary comparison of models
tab_model(mod.foliarlitter1, mod.foliarlitter2, mod.foliarlitter3)
tab_model(mod.littersoil1, mod.littersoil2, mod.littersoil3)
tab_model(mod.foliarsoil1, mod.foliarsoil2, mod.foliarsoil3)


#visualize best model from each (based on marginal and conditional R2)
sjPlot::plot_model(mod.foliarlitter3, 
                   #axis.labels=c("Urchin", "Depth", "Fish"),
                   show.values=TRUE, show.p=TRUE,
                   title="Effect of Parameters on difference in CN Ratio (foliar - litter")
#visualize best model from each (based on marginal and conditional R2)
sjPlot::plot_model(mod.littersoil3, 
                   #axis.labels=c("Urchin", "Depth", "Fish"),
                   show.values=TRUE, show.p=TRUE,
                   title="Effect of Parameters on difference in CN Ratio (litter - soil)")
#visualize best model from each (based on marginal and conditional R2)
sjPlot::plot_model(mod.foliarsoil3, 
                   #axis.labels=c("Urchin", "Depth", "Fish"),
                   show.values=TRUE, show.p=TRUE,
                   title="Effect of Parameters on difference CN Ratio (foliar - soil)")


```

## Using CN as a predictor in models
Re-run Andrew's models with delta CN to see how they change. Overall, model accuracy is increased when either CN or delta CN is included:
```{r cn models}
mod.soil0 <- lmer(soilNPercent_MHoriz ~ litterNPercent + foliarNPercent + (1|siteID), data=plot.df)
mod.soil1 <- lmer(soilNPercent_MHoriz ~ litterCNratio + soilCNRatio_MHoriz + litterNPercent + foliarNPercent + (1|siteID), data=plot.df)

mod.litter0 <- lmer(litterNPercent ~ litterNPercent + soilNPercent_MHoriz + (1|siteID), data=plot.df)
mod.litter1 <- lmer(litterNPercent ~ litterCNratio + soilCNRatio_MHoriz + soilNPercent_MHoriz + foliarNPercent + (1|siteID), data=plot.df)

mod.foliar0 <- lmer(foliarNPercent ~ litterNPercent + foliarNPercent + (1|siteID), data=plot.df)
mod.foliar1 <- lmer(foliarNPercent ~ litterCNratio + soilCNRatio_MHoriz + litterNPercent + soilNPercent_MHoriz + (1|siteID), data=plot.df)



tab_model(mod.soil0, mod.soil1)
tab_model(mod.litter0, mod.litter1)
tab_model(mod.foliar0, mod.foliar1)
```
